---
title       : Find Universities by SAT Score
subtitle    : Using data published by College Scorecard
author      : Rupendra Bandyopadhyay
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

US Department of Education publishes data for US Universities that presents SAT scores for 25th, 50th and 75th percentile of admitted students by institute. This data is available at the [College Scorecard Website.](https://collegescorecard.ed.gov/data)

Given this information, students should be able to find out which Universities typically admit students with scores close to theirs. Since we only have data for three discrete percentiles per University for each of the three SAT scores (reading, writing and math), we can only match a student to Universities given their preference of one of those specific percentiles of admitted students they would like to belong to.

So, an automated system would need the student to enter her SAT scores (for Reading, Writing and Math) and her preferred percentile rank (choice of 25, 50 or 75) by SAT scores in the institute where she would like to apply.

With this information it is possible determine the closeness of a college to a student by simply summing up the absolute differences between her scores and the universities' published scores. This closeness score can be used to display a list of matching institutes.

```{r setOptions, echo=FALSE}
options(plot.gvis.tag="chart")
options(gvis.plot.tag="chart")
```

--- .class #id 

## Determine match and present results

<b>The application accepts the following inputs from user:</b>  

1. SAT Reading, Writing and Math scores  (Actual or projected scores for scenario analysis)
2. Preferred percentile rank in the institution  (25, 50 or 70, defaults to 50)
3. Number of closest matching institutions to show  

<b>The application does the following:</b>  

1. Calculates a score difference for each of the universities by summing up the absolute deviations from the input values  
2. Filters the list by institutes with score difference <= 60  
3. Orders the list by decreasing score difference and picks top n based on input  
4. Shows the results in a tabular form  
5. Shows the results on a map with both size and color gradient emphasizing match proximity  




--- 

## How it works (the core code)

```{r echo=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(googleVis))
```

```{r results='asis'}
source("resources.R") # Source code for needed R functions
newData <- download_read_and_cleanse_scorecard_file()
satWrite <- 670; satRead <- 650; satMath <- 700; preferredPctRank <- 50; topn <- 10
ret <- getUnivMatches(satWrite, satMath, satRead, preferredPctRank, topn)
mychart <- ret("chartonly")
plot(ret("dataonly"))
```

--- 

## And a Plot

```{r results='asis'}
plot(mychart)
```
### Please visit the website at [https://rupenb.shinyapps.io/univsearch](https://rupenb.shinyapps.io/univsearch)
---

